Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-10-01 paper.qmd
Updated Figure 10

distance_two_curvy_dist <- distance_two_curvy_dist |>
  count(bin) |>
  mutate(prop = n/sum(n))

@@

distance_hist <- ggplot(
  distance_two_curvy_dist, 
  aes(x = as.numeric(bin),
      y = prop)) +
  geom_line() +
  scale_x_continuous(breaks = as.numeric(distance_two_curvy_dist$bin),
                     labels = c(
                       expression(a[1]),
                       expression(2*a[1]),
                       expression(3*a[1]),
                       expression(4*a[1]),
                       expression(5*a[1]),
                       expression(6*a[1]),
                       expression(7*a[1]),
                       expression(8*a[1])
                     )) +
  geom_vline(xintercept = 2,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 4.5,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  ylab("Proportion of the number of edges") +
  xlab(expression(d^{(2)})) +
  #ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

benchmark_two_curvy <- 4.5 * bin_width

@@

free(distance_hist) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1,
two_curvy_proj_umap_model_delaunay,
two_curvy_proj_umap_model_benchmark2,
two_curvy_proj_umap_model, ncol = 3) 
```

>>>>

distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))

distance_points <- ggplot(
  distance_two_curvy_dist, 
  aes(x = as.numeric(bin),
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  geom_jitter() +
  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
                     labels = c(
                       expression(a[1]),
                       expression(2*a[1]),
                       expression(3*a[1]),
                       expression(4*a[1]),
                       expression(5*a[1]),
                       expression(6*a[1]),
                       expression(7*a[1]),
                       expression(8*a[1])
                     )) +
  # geom_vline(xintercept = 2,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  # geom_vline(xintercept = 4.5,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  #ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

distance_two_curvy_dist_summary <- distance_two_curvy_dist |>
  count(bin) |>
  mutate(prop = n/sum(n)) |>
  head(4)

@@

distance_hist <- ggplot(
  distance_two_curvy_dist_summary,
  aes(x = as.numeric(bin),
      y = prop)) +
  geom_line() +
  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin)))[1:4],
                     labels = c(
                       expression(a[1]),
                       expression(2*a[1]),
                       expression(3*a[1]),
                       expression(4*a[1])
                     )) +
  geom_vline(xintercept = 2,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 3.5,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  ylab("Proportion of the number of edges") +
  xlab(expression(d^{(2)})) +
  #ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

@@

benchmark_two_curvy <- 3.5 * bin_width

@@

wrap_plots(distance_points,distance_hist, ncol = 2) / wrap_plots(trimesh_two_curvy2, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1,
two_curvy_proj_umap_model_delaunay,
two_curvy_proj_umap_model_benchmark2,
two_curvy_proj_umap_model, ncol = 3) 
```
